Meta-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates

### 2025-01-27 paper/1_Initial/manuscript.tex
minor changes

\authorsnames{Ana Neves,Magdalena Pfaff,Dominique Makowski}

@@

\leftheader{Neves, Pfaff and Makowski}

@@

\authornote{\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Magdalena
Pfaff}{0009-0006-2386-7936}}\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}

\par{     \begin{tcolorbox}[enhanced jigsaw, colback=white, breakable, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, arc=.35mm, toprule=.15mm, rightrule=.15mm, opacityback=0, left=2mm, leftrule=.75mm]

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\begin{center}
\includegraphics[width=0.2\textwidth,height=\textheight]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
\end{center}

\end{tcolorbox}  }
\par{Correspondence concerning this article should be addressed
to Dominique Makowski, Email: D.Makowski@sussex.ac.uk}
}

@@

\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

{[}\textbf{TO DO}: add - previous work suggests the importance of
physiological contexts (Vlemincx et al., 2021){]}

@@

For instance, while the use of self-reports questionnaires to measure an
embodied function might seem paradoxical, recent redefinitions of
interoception, emphasizing the role of high-level and metacognitive
elaboration of interoceptive information, has provided the theoretical
grounding to support the idea that some facets of interoception
(including participants' metacognitive beliefs) can be assessed
subjectively, providing useful and interesting measures
(\citeproc{ref-lin2023}{Lin et al., 2023};
\citeproc{ref-murphy2019}{Murphy et al., 2019}). The notion that
self-reports might not reflect the same processes as other interoception
measures is important to contextualize the apparent lack of convergence
between measures in the field
(\citeproc{ref-desmedt2022measures}{Desmedt et al., 2022}). Still, a
better understanding of what is being measured with different
questionnaires and dimensions, as well their potential overlaps with
other constructs (e.g., alexithymia, body awareness) is needed to
clarify the role of self-reports in the assessment of interoception.

@@

Although the original validation study suggested a two-factor solution
structure, the authors underlines the need for further investigation of
the scale's factor structure, as it provided acceptable but imperfect
explanation of the data {[}Murphy et al.
(\citeproc{ref-murphy2019}{2019}); p.~127{]}. Notably, the only other
validation study to report a 2-factor solution was conducted by Koike
and Nomura (\citeproc{ref-koike2023}{2023}), who performed an
Exploratory Factor Analysis (EFA) assuming 2 factors to align with the
findings from the original validation paper. Other follow-up studies
using confirmatory factor analysis (CFA) and structural modeling
(\citeproc{ref-morin2016}{Morin et al., 2015}) have identified different
optimal solutions. Some studies, like Brand et al.
(\citeproc{ref-brand2023}{2023}), reported a 1-factor solution, while
Lin et al. (\citeproc{ref-lin2023}{2023}) and Campos et al.
(\citeproc{ref-campos2021}{2021}) found bifactor solutions - one general
factor and a set of loweroike -level factors (Rodriguez et al.
(\citeproc{ref-rodriguez2016evaluating}{2016})) - to be the best fit.

Discussions have also been focused on specific items. For instance,
Murphy et al. (\citeproc{ref-murphy2019}{2019}) notes that some items
might measure direct interoceptive signals such as cardioception, while
others might capture phenomena not perceivable through interoceptive
signals alone (e.g., ``bruising''; p.~119). Lin et al.
(\citeproc{ref-lin2023}{2023}) also highlights their correlation
analysis, showing five locally dependent pairs and three items (touch,
blood sugar, bruise) with exceptionally high difficulty and low
discrimination. Additionally, Campos et al.
(\citeproc{ref-campos2021}{2021}) suggests that the ``tickle'' item
represents a more specific factor. Interestingly, Lin et al.
(\citeproc{ref-lin2023}{2023}) reported that all items of the IAS
grouped together using a new approach, Exploratory Graph Analysis
{[}EGA; Golino and Epskamp
(\citeproc{ref-golino2017exploratory}{2017}){]}, to assess convergent
and discriminant validity, which suggests unidimensionality.

While the assessment of the validity of an interoception scale can be
conceived as theoretically challenging, several traits have been used to
assess convergent validity for the the IAS, including expected negative
associations with alexithymia Murphy et al.
(\citeproc{ref-murphy2019}{2019}), somatic symptoms Lin et al.
(\citeproc{ref-lin2023}{2023}), depressive symptoms
(\citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023};
\citeproc{ref-lin2023}{Lin et al., 2023}), anxiety
(\citeproc{ref-brand2023}{Brand et al., 2023}), neuroticism
(\citeproc{ref-brand2023}{Brand et al., 2023}) and self-esteem
(\citeproc{ref-murphy2019}{Murphy et al., 2019}).

The IAS has also been compared to other interoception scales (and
interoceptive related scales),generally showing positive correlations
with most facets of the MAIA, except for the Non-Distracting and
Not-Worrying subscales (\citeproc{ref-brand2023}{Brand et al., 2023}).
Findings on the correlation between the IAS and the body awareness
dimension of the Body Perception Questionnaire (BPQ-A; Porges
(\citeproc{ref-porges1993body}{1993})) have been mixed: some studies
report small positive correlations (\citeproc{ref-brand2023}{Brand et
al., 2023}; \citeproc{ref-campos2021}{Campos et al., 2021};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023}), while others find
small negative correlations (\citeproc{ref-lin2023}{Lin et al., 2023})
or no correlation at all (\citeproc{ref-murphy2019}{Murphy et al.,
2019}). Small positive correlations have also been observed with the
observation and description subscales of the Five Facet Mindfulness
Questionnaire (FFMQ, Baer et al. (\citeproc{ref-baer2006using}{2006});
Brand et al. (\citeproc{ref-brand2023}{2023}); Koike and Nomura
(\citeproc{ref-koike2023}{2023})), as well as with the nonreactivity and
acting with awareness subscales (\citeproc{ref-koike2023}{Koike \&
Nomura, 2023}). Additionally, the IAS has shown a positive correlation
with the interoceptive awareness subscale of the Eating Disorder
Inventory (\citeproc{ref-lin2023}{Lin et al., 2023}) and a negative
correlation with the Interoceptive Confusion Questionnaire
(\citeproc{ref-brand2023}{Brand et al., 2023}; ICQ,
\citeproc{ref-brewer2016alexithymia}{Brewer et al., 2016};
\citeproc{ref-murphy2019}{Murphy et al., 2019}). Lastly, small positive
correlations have been reported between the Interoceptive Attention
Scale (IATS, Gabriele et al.
(\citeproc{ref-gabriele2022dissociations}{2022})) and the IAS
(\citeproc{ref-koike2023}{Koike \& Nomura, 2023};
\citeproc{ref-lin2023}{Lin et al., 2023}), though studies have also
found no correlation between these measures
(\citeproc{ref-gabriele2022dissociations}{Gabriele et al., 2022}).

@@

The goal of study 1 is to re-analyse and assess the factor structure of
the IAS by taking advantage of the large number of open-access datasets
(\citeproc{ref-arslanova2022}{Arslanova et al., 2022};
\citeproc{ref-brand2022}{Brand et al., 2022};
\citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-campos2021}{Campos et al., 2021};
\citeproc{ref-gaggero2021}{Gaggero et al., 2021};
\citeproc{ref-lin2023}{Lin et al., 2023};
\citeproc{ref-murphy2019}{Murphy et al., 2019};
\citeproc{ref-todd2022}{Todd et al., 2022}; \citeproc{ref-von2023}{Von
Mohr et al., 2023}). While combining these studies might provide a more
robust and generalizable understanding of the IAS' factor structure, it
is important to note that the studies differ in their sample sizes,
demographics, language, and the way they collected the data. Thus,
providing additionally an individual analysis (i.e., on all samples
individually) will help adding nuance to the general picture.

@@

\input{../../analysis/figures/table1.tex}

@@

\input{../../analysis/figures/table2.tex}

@@

@@

@@

@@

\bibitem[\citeproctext]{ref-morin2016}
Morin, A. J. S., Arens, A. K., \& Marsh, H. W. (2015). A bifactor
exploratory structural equation modeling framework for the
identification of distinct sources of construct-relevant psychometric
multidimensionality. \emph{Structural Equation Modeling: A
Multidisciplinary Journal}, \emph{23}(1), 116--139.
\url{https://doi.org/10.1080/10705511.2014.961800}

>>>>

\author{Ana Neves}

@@

\leftheader{Neves}

@@

\authornote{\par{\addORCIDlink{Ana Neves}{0009-0006-0020-7599}}

\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, bottomrule=.15mm, rightrule=.15mm, toprule=.15mm, left=2mm, colback=white, arc=.35mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm, opacityback=0]

\end{tcolorbox}

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\begin{center}
\includegraphics[width=0.2\textwidth,height=\textheight]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
\end{center}  Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Ana
Neves:   Project administration, Data curation, Formal
Analysis, Investigation, Visualization, Writing -- original
draft, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed to }
}

@@

\usepackage{lscape}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

@@

The use of subjective self-report questionnaires to measure deeply
embodied functions might seem paradoxical. However, recent redefinitions
of interoception emphasize the role of high-level and metacognitive
elaboration of interoceptive information. These redefinitions provide
theoretical grounding to support the idea that some facets of
interoception, including participants' metacognitive beliefs, can be
assessed subjectively. This approach offers useful and interesting
measures (\citeproc{ref-lin2023}{Lin et al., 2023};
\citeproc{ref-murphy2019}{Murphy et al., 2019}).

The notion that self-reports might not reflect the same processes as
other interoception measures is important to contextualize the apparent
lack of convergence between measures in the field
(\citeproc{ref-desmedt2022measures}{Desmedt et al., 2022}).

A recent systematic review by Desmedt et al.
(\citeproc{ref-desmedt2022measures}{2022}) examined whether various
questionnaires designed to measure ``interoceptive sensibility'' truly
assess the same construct. Notably, Garfinkel et al.
(\citeproc{ref-garfinkel2015knowing}{2015}) defines interoceptive
sensibility as the self-reported tendency to focus on and detect
internal sensations, whereas
(\citeproc{ref-khalsa2018interoception}{\textbf{khalsa2018interoception?}})
defines it more narrowly, excluding the ability to detect these signals.
According to Desmedt et al. (\citeproc{ref-desmedt2022measures}{2022}),
most authors adopt the latter definition. The review found that these
questionnaires measure related but distinct constructs, leading
researchers to commit the `jingle fallacy'---mistakenly treating them as
equivalent measures of interoceptive sensibility.

\textbf{TODO: do a better job}

A better understanding of what is being measured with different
questionnaires and dimensions, as well as their potential overlaps with
other constructs (e.g., alexithymia, body awareness), is thus needed to
clarify the role of self-reports in the assessment of interoception.

@@

Although the original validation study suggested a two-factor structure
for the IAS, the authors underline its acceptable but imperfect fit
{[}Murphy et al. (\citeproc{ref-murphy2019}{2019}); p.~127{]}, calling
on further investigation of the scale's factor structure. Notably, the
authors did not define these factors, as no clear explanations were
evident. They suggested that the first factor reflects the perception of
interoceptive signals, while the second pertains to signals that may be
difficult to perceive solely through interoceptive information.

Other follow-up studies using confirmatory factor analysis (CFA) and
structural modeling have identified different optimal solutions. Some
studies, like Brand et al. (\citeproc{ref-brand2023}{2023}), reported a
1-factor solution, while Lin et al. (\citeproc{ref-lin2023}{2023}) and
Campos et al. (\citeproc{ref-campos2021}{2021}) found bifactor solutions
- one general factor and a set of lower-level factors
(\citeproc{ref-rodriguez2016evaluating}{Rodriguez et al., 2016}) - to be
the best fit. Notably, the only other study to report a 2-factor
solution was conducted by Koike and Nomura
(\citeproc{ref-koike2023}{2023}), who performed an Exploratory Factor
Analysis (EFA) assuming 2 factors to align with the findings from the
original validation paper.

Discussions have also been focused on specific items. For instance,
Murphy et al. (\citeproc{ref-murphy2019}{2019}) notes that some items
might measure direct interoceptive signals such as cardioception, while
others might capture phenomena not perceivable through interoceptive
signals alone (e.g., ``bruising''; p.~119). Lin et al.
(\citeproc{ref-lin2023}{2023}) also highlights their correlation
analysis, showing five locally dependent pairs and three items (touch,
blood sugar, bruise) with exceptionally high difficulty and low
discrimination. Additionally, Campos et al.
(\citeproc{ref-campos2021}{2021}) reported ``tickle'' to be the only
item that reflected more specific factors than the general factor.
Interestingly, Lin et al. (\citeproc{ref-lin2023}{2023}) reported that
all items of the IAS grouped together using a new approach, Exploratory
Graph Analysis {[}EGA; Golino and Epskamp
(\citeproc{ref-golino2017exploratory}{2017}){]}, to assess convergent
and discriminant validity, providing further evidence for
unidimensionality.

The IAS has naturally been compared to other interoception-related
scales, and shows a positive correlations with most facets of the MAIA
(\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018}),
except for the Non-Distracting and Not-Worrying subscales
(\citeproc{ref-brand2023}{Brand et al., 2023}). Interestingly, findings
on the correlation between the IAS and the body awareness dimension of
the Body Perception Questionnaire (BPQ-A,
\citeproc{ref-porges1993body}{Porges, 1993}) have been mixed: some
studies report small positive correlations
(\citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-campos2021}{Campos et al., 2021};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023}), while others find
small negative correlations (\citeproc{ref-lin2023}{Lin et al., 2023})
or no correlation at all (\citeproc{ref-murphy2019}{Murphy et al.,
2019}). Small positive correlations have also been observed with the
``observation'' and ``description'' subscales of the Five Facet
Mindfulness Questionnaire (FFMQ, \citeproc{ref-baer2006using}{Baer et
al., 2006}; \citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023}), as well as with the
``non-reactivity'' and ``acting with awareness'' subscales
(\citeproc{ref-koike2023}{Koike \& Nomura, 2023}). Additionally, the IAS
has shown a positive correlation with the interoceptive awareness
subscale of the Eating Disorder Inventory (\citeproc{ref-lin2023}{Lin et
al., 2023}) and a negative correlation with the Interoceptive Confusion
Questionnaire (\citeproc{ref-brand2023}{Brand et al., 2023}; ICQ,
\citeproc{ref-brewer2016alexithymia}{Brewer et al., 2016};
\citeproc{ref-murphy2019}{Murphy et al., 2019}). Lastly, small positive
correlations have also been reported with the Interoceptive Attention
Scale (\citeproc{ref-koike2023}{Koike \& Nomura, 2023}; IATS,
\citeproc{ref-lin2023}{Lin et al., 2023}), though studies have also
found no correlation between these measures
(\citeproc{ref-gabriele2022dissociations}{Gabriele et al., 2022}).

While assessing the validity of an interoception scale can be conceived
as theoretically challenging, several measures have been used to assess
convergent validity for the the IAS, including expected negative
associations with alexithymia Murphy et al.
(\citeproc{ref-murphy2019}{2019}), somatic symptoms Lin et al.
(\citeproc{ref-lin2023}{2023}), depressive symptoms
(\citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023};
\citeproc{ref-lin2023}{Lin et al., 2023}), anxiety
(\citeproc{ref-brand2023}{Brand et al., 2023}), neuroticism
(\citeproc{ref-brand2023}{Brand et al., 2023}) and self-esteem
(\citeproc{ref-murphy2019}{Murphy et al., 2019}).

@@

The goal of study 1 is to re-analyse and assess the factor structure of
the IAS by taking advantage of the large number of open-access datasets
(\citeproc{ref-arslanova2022}{Arslanova et al., 2022};
\citeproc{ref-brand2022}{Brand et al., 2022};
\citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-campos2021}{Campos et al., 2021};
\citeproc{ref-gaggero2021}{Gaggero et al., 2021};
\citeproc{ref-lin2023}{Lin et al., 2023};
\citeproc{ref-murphy2019}{Murphy et al., 2019};
\citeproc{ref-todd2022}{Todd et al., 2022}; \citeproc{ref-von2023}{Von
Mohr et al., 2023}). While combining these studies might provide a more
robust and generalizable understanding of the IAS' factor structure, we
also additionally provide an individual analysis (i.e., on all samples
separately) to add nuance to the general picture, as all studies differ
in their sample sizes, demographics, language, and procedure.

@@

@@

@@

\textbf{todo: compute average correlations }

@@

{[}\textbf{TO DO}: add - previous work suggests the importance of
physiological contexts (Vlemincx et al., 2021){]} \textbf{I would rather
put that in the discussion in the suggestions for better scales}

@@

\bibitem[\citeproctext]{ref-garfinkel2015knowing}
Garfinkel, S. N., Seth, A. K., Barrett, A. B., Suzuki, K., \& Critchley,
H. D. (2015). Knowing your own heart: Distinguishing interoceptive
accuracy from interoceptive awareness. \emph{Biological Psychology},
\emph{104}, 65--74.

@@
